\section{Deterministic Simulation Testing (DST)-style fault injection}
\label{sec:dst}

We extend Cogitator with deterministic fault injection (DST-style chaos) as a controlled
experimental variable. The fault schedule is seeded and keyed by run identifiers,
step indices, tool call indices, and domain keys, yielding a reproducible fault decision
and parameters (e.g., corruption masks). Fault selection uses a single deterministic
draw with cumulative per-million weights in fixed priority order (timeout, drop, corrupt,
latency\_sim), with the total fault probability capped at 1{,}000{,}000. This schedule is
written to a
\texttt{chaos\_profile.json} artifact and its summary is committed into the witness
metadata, ensuring the witness root is invariant for identical inputs, seed, transcript,
and fault schedule.

Faults are injected at the tool boundary (timeouts, payload corruption, drops, and
simulated latency). Simulated latency is exposed to the agent but excluded from witness
commitments unless explicitly configured, maintaining auditability without binding to
wall-clock timing. Canonical JSON serialization (RFC~8785 JCS) is used across artifacts
to ensure byte-stable outputs for the same declared inputs.

\section{Commitment boundaries}
\label{sec:commitments}

Cogitator separates witnessed commitments from provenance. The witness root commits to
canonical trace events plus the agent trace and tool-call witness view (agent step, then
tool calls ordered by \texttt{tool\_call\_idx}). Runtime context such as wall-clock time,
thread counts, and environment fingerprints are captured only in provenance and never
enter witnessed commitments. The witness bundle hash additionally commits to all emitted
artifacts (including optional provenance artifacts like \texttt{nix\_provenance.json})
for offline verification.

LLM inference is modeled as a tool call: in live mode, responses are recorded into the
tool transcript, and in replay mode those responses are replayed verbatim to preserve
determinism. The default backend is a deterministic stub, so no model installation or
network access is required unless explicitly enabled.

\section{Related work}
\label{sec:related}

Cogitator's deterministic chaos testing parallels the simulation framework for
FoundationDB, which treats fault injection as a first-class, reproducible experimental
variable. Reproducibility as an evaluated property is emphasized in AIRepr, which
provides structured metadata and audit trails for AI experiments. Canonical JSON
serialization follows RFC~8785 (JSON Canonicalization Scheme) to ensure stable artifact
hashes.

\begin{thebibliography}{9}
\bibitem{foundationdb_simulation}
Peter Bailis et al. \emph{FoundationDB Simulation Testing}. FoundationDB Engineering,
2014. Available at \url{https://apple.github.io/foundationdb/simulation.html}.

\bibitem{airepr}
Ying Jiang et al. \emph{AIRepr: A Reproducibility Framework for AI Experiments}. 2023.
Available at \url{https://arxiv.org/abs/2303.01555}.

\bibitem{rfc8785}
S. Hassas Yeganeh, Y. Ganjali. \emph{RFC 8785: JSON Canonicalization Scheme (JCS)}.
IETF, 2020. Available at \url{https://www.rfc-editor.org/rfc/rfc8785}.
\end{thebibliography}
